---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->
[![R-CMD-check](https://github.com/benet1one/lpsugar/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/benet1one/lpsugar/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

# Linear, Quadratic and Nonlinear Programming in R

`lpsugar` is an Algebraic Modeling Language (AML) inspired in R’s
vectorized syntax, used to define and solve Mixed Integer Linear
Problems (MILP), Mixed Integer Quadratic Problems (MIQP) and
General Nonlinear Problems (NLP).

This package is in development, and being made by a solo student. 
For a safer choice, I recommend using [ompr](https://dirkschumacher.github.io/ompr/).

## Installation

You can install the development version of `lpsugar` from 
[GitHub](https://github.com/benet1one/lpsugar) with:

```r
# install.packages("pak")
pak::pak("benet1one/lpsugar")
```

## Quick Examples

Start with a simple problem to show the basic syntax.

$$
\begin{array}{rl}
  \text{max} & x+y \\
  \text{st} & 2x + y \le 8 \\
            & 2x + 3y \le 12
\end{array}
$$

Let's write the problem in `lpsugar` and solve it!

```{r example}
library(lpsugar)
library(ROI.plugin.highs) # Use any ROI solver you like!

my_problem <- lp_problem() |> 
    lp_variable(x) |> 
    lp_variable(y) |> 
    lp_maximize(x + y) |> 
    lp_constraint(
        2*x + y <= 8,
        2*x + 3*y <= 12
    )

my_solution <- lp_solve(my_problem)
my_solution$variables
my_solution$objective
```

### Transportation Problem

Here's an example solving the classic transportation problem. We shall transport
`x[f, m]` units from factory `f` to market `m`, subject to a `supply[f]`
and a `demand[m]`. We will minimize the total transportation `cost`.

```{r}
supply <- c(BLB = 20, ACO = 25, GRN = 15)
demand <- c(Madrid = 8, Barcelona = 10, Valencia = 12, Seville = 10)

factory <- names(supply)
market <- names(demand)

cost <- c(
    2, 7, 3, 1,
    6, 2, 9, 2,
    4, 6, 2, 8
) |> parameter(factory, market)

cost

transportation_problem <- lp_problem() |> 
    lp_variable(
        x[factory, market], 
        integer = TRUE, 
        lower = 0
    ) |> 
    lp_minimize(
        sum_over(f = factory, m = market, x[f, m] * cost[f, m])
    ) |> 
    lp_constraint(
        for (f in factory) sum(x[f, ]) <= supply[f],
        for (m in market)  sum(x[, m]) >= demand[m]
    )

transportation_solution <- lp_solve(transportation_problem)
transportation_solution$variables$x
transportation_solution$objective
```

## Vignettes

For more examples, see the [vignettes](https://benet1one.github.io/lpsugar/articles/)
in the package website.
