---
title: "browsers"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{browsers}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(lpsugar)
```

In this vignette we'll optimize a set of parameters to prove that our
browser is the best browser.

## The Problem

Imagine we're the developers of Arc browser, and we want to prove that it's
better than other any other browser. We have a table of features for each of the browsers,
that we can quickly make up.

```{r}
browsers <- c("Arc", "Chrome", "Firefox", "Opera", "Safari")
features <- c("Speed", "Ease of use", "Customizability", "Privacy", "Extensions",
              "Reading mode", "Ad-blocker")

# Make up ratings from 0 to 10 for each feature
set.seed(8456)
ratings <- parameter(0, features, browsers)
ratings[] <- sample(0:10, size = length(ratings), replace = TRUE)

ratings
```

The question is: how important is each feature? Let's try giving all of them
the same weight and calculate the score each browser gets.

```{r}
colSums(ratings)
```

Nonsense, this would mean Arc is worse than Opera or Safari, and much worse than Firefox! 
Let's give each feature $i$ an optimal weight $w_i$. Then we maximize the gap between Arc's score
and the second best score.

$$
\max_w {\text{ArcScore}(w) - \text{SecondBestScore}(w)}
$$

We'll make sure each feature *matters* by having $1 < w_i < 5$, and make
them integers for simplicity.

```{r}
browser_problem <- lp_problem() |> 
    lp_variable(weights[features], lower = 1, upper = 5, integer = TRUE) |> 
    lp_variable(arc_score) |> 
    lp_variable(second_best) |>
    lp_maximize(arc_score - second_best) |> 
    lp_subject_to(
        arc = {
            arc_ratings <- ratings[, "Arc"]
            arc_score == sum(arc_ratings * weights)
        },
        others = for (b in browsers[browsers != "Arc"]){
            b_ratings <- ratings[, b]
            second_best >= sum(b_ratings * weights)
        }
    )

library(ROI.plugin.highs)
browser_solution <- lp_solve(browser_problem)
browser_solution$variables
```

Who cares about speed anyway? Extensions are where it's at. Take a look at the new scores:

```{r}
optimal_weights <- browser_solution$variables$weights
apply(ratings, 2, \(r) sum(r * optimal_weights))
```

That's much better. We have successfully proved Arc's unquestionable superiority!
