---
title: "gerrymandering"
output: rmarkdown::html_vignette
vignette: >
%\VignetteIndexEntry{gerrymandering}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  ---
  
  ```{r, include = FALSE}
knitr::opts_chunk$set(
collapse = TRUE,
comment = "#>"
)
```

```{r setup}
library(lpsugar)
library(ROI.plugin.highs)

library(dplyr, warn.conflicts = FALSE)
library(ggplot2)
```

## Data

Generating Random Regions

```{r}
set.seed(155)
rows <- 6
cols <- 10

regions <- expand.grid(i = 1:rows, j = 1:cols) |> 
    as_tibble() |> 
    mutate(
        ind = 1:n(),
        population = round(rexp(n(), 1/500)),
        teal_prop = runif(n(), 0.15, 0.95),
        teal_votes = round(population * teal_prop),
        gold_votes = population - teal_votes
    )

votes_plot <- ggplot(regions, aes(x = j, y = i, color = teal_prop, size = population)) +
    geom_point() +
    scale_size(name = "Population", range = c(6, 14)) +
    scale_color_steps(
        name = "Proportion of Teal Votes", 
        n.breaks = 5, limits = c(0, 1),
        low = "#FADF63", high = "#48A9A6"
    ) +
    theme_void()

votes_plot
```

Function to get indices of contiguous regions

```{r}
contiguous <- function(region_ind) {
    ri <- regions$i[region_ind]
    rj <- regions$j[region_ind]
    
    cont <- regions |> filter(
        abs(i - ri) <= 1,
        abs(j - rj) <= 1,
        abs(i - ri) + abs(j - rj) == 1
    )
    
    cont$ind
}

network <- tibble(from = regions$ind) |> 
    group_by(from) |> 
    reframe(to = contiguous(from)) |> 
    mutate(ind = 1:n(), .before = 1)
```

## Problem

Given a desired number of districts and a minimum number of teal districts (where
the teal party is expected to win), maximizes the smallest margin of victory across
all teal districts. Districts will have at least `min_district_regions` and, 
at most, `max_district_regions`.

### Parameters

```{r}
n_districts = 4
min_teal_districts = 3
min_district_regions = 2
max_district_regions = floor(0.6 * nrow(regions))
```

### Model

```{r}
R <- regions$ind
L <- network$ind
D <- LETTERS[1:n_districts]
DT <- D[1:min_teal_districts]

# set.seed(123)
which_source <- parameter(0, D)
which_source[1] <- which.max(regions$teal_prop)
which_source[n_districts] <- which.min(regions$teal_prop)
which_source[2:(n_districts - 1)] <- sample.int(nrow(regions), size = 2)
source <- parameter(0, R, D)

for (d in D) {
    r <- which_source[d]
    source[r, d] <- 1
}

p <- lp_problem() |> 
    lp_var(x[R, D], binary = TRUE) |> 
    lp_var(smallest_victory_margin) |>
    lp_max(smallest_victory_margin) |> 
    
    # lp_var(source[R, D], binary = TRUE) |>
    lp_var(flow[L, D], lower = 0) |> 
    
    lp_con(
        one_district_per_region = for (r in R) {
            sum(x[r, ]) == 1
        },
        regions_per_district = for (d in D) bind_cons(
            sum(x[, d]) >= min_district_regions,
            sum(x[, d]) <= max_district_regions
        ),
        victory_margins = for (d in DT) {
            teal_votes_d <- sum_over(r = R, regions$teal_votes[r] * x[r, d])
            gold_votes_d <- sum_over(r = R, regions$gold_votes[r] * x[r, d])
            teal_votes_d >= gold_votes_d + smallest_victory_margin
        },
        
        # one_source_per_district = for (d in D) {
        #     sum(source[, d]) == 1
        # },
        # sources_per_region = for (r in R) {
        #     sum(source[r, ]) <= 1
        # },
        
        source_must_belong_to_district = {
            source <= x
        },
        flow_demand = for (d in D) for (r in R) {
            BigM <- 2*nrow(regions)
            supply <- BigM * source[r, d]
            
            links_in <- network |> filter(to == r) |> _$ind
            links_out <- network |> filter(from == r) |> _$ind
            
            flow_in <- sum(flow[links_in, d])
            flow_out <- sum(flow[links_out, d])
            
            bind_cons(
                flow_in <= BigM * x[r, d],
                flow_out <= BigM * x[r, d],
                x[r, d] <= supply + flow_in - flow_out
            )
        }
    )
```

### Solution

```{r echo=FALSE}
s <- lp_solve(
    p, 
    binary_as_logical = TRUE, 
    time_limit = 60,
    log_to_console = TRUE
)
print(s$status)
```

### Analysis

```{r}
which_district <- apply(s$variables$x, 1, which)

districts <- regions |> 
    mutate(district = factor(D[which_district])) |> 
    print()

district_summary <- districts |> 
    group_by(district) |> 
    summarise(
        dt_population = sum(population),
        dt_teal_votes = sum(teal_votes),
        dt_gold_votes = sum(gold_votes),
        dt_teal_prop = dt_teal_votes / dt_population,
        dt_diff = dt_teal_votes - dt_gold_votes,
    )

district_summary
```

### Plot

```{r fig.height=6}
plot_districts <- function(districts) {
    network2 <- network |> 
        mutate(
            i_from = regions$i[from],
            j_from = regions$j[from],
            i_to = regions$i[to],
            j_to = regions$j[to],
            horizontal = i_from == i_to,
            
            dt_from = districts$district[from],
            dt_to = districts$district[to],
            same_district = dt_from == dt_to,
        )
    
    network_h <- network2 |> 
        filter(horizontal, !same_district) |> 
        mutate(
            x = (j_from + j_to) / 2,
            xend = x,
            y = i_from - 0.5,
            yend = i_from + 0.5
        )
    
    network_v <- network2 |>
        filter(!horizontal, !same_district) |> 
        mutate(
            x = j_from - 0.5,
            xend = j_from + 0.5,
            y = (i_from + i_to) / 2,
            yend = y
        )
    
    network3 <- bind_rows(network_h, network_v)
    
    votes_plot +
        geom_raster(
            data = districts, 
            aes(fill = district), 
            alpha = 0.08, show.legend = FALSE
        ) +
        geom_segment(
            data = network3, 
            inherit.aes = FALSE,
            aes(x = x, xend = xend, y = y, yend = yend),
            linewidth = 1.5, lineend = "round", color = "gray30"
        ) +
        scale_fill_viridis_d(option = "A")
}

plot_districts(districts)
```

## Reversing the Problem

Let us now redistrict in Gold's favour.

```{r}
min_gold_districts = 3
DG <- D[1:min_gold_districts]

p_rev <- p |> 
    lp_delete_constraint("victory_margins") |> 
    lp_constraint(
        victory_margins_rev = for (d in DT) {
            teal_votes_d <- sum_over(r = R, regions$teal_votes[r] * x[r, d])
            gold_votes_d <- sum_over(r = R, regions$gold_votes[r] * x[r, d])
            gold_votes_d >= teal_votes_d + smallest_victory_margin
        }
    )
```

### New Solution

```{r}
s_rev <- lp_solve(
    p_rev,
    binary_as_logical = TRUE,
    time_limit = 60,
    log_to_console = TRUE
)

print(s_rev$status)
```

### Analysis

```{r}
which_district <- apply(s_rev$variables$x, 1, which)

districts_rev <- regions |> 
    mutate(district = factor(D[which_district])) |> 
    print()

district_summary_rev <- districts_rev |> 
    group_by(district) |> 
    summarise(
        dt_population = sum(population),
        dt_teal_votes = sum(teal_votes),
        dt_gold_votes = sum(gold_votes),
        dt_teal_prop = dt_teal_votes / dt_population,
        dt_diff = dt_teal_votes - dt_gold_votes,
    )

district_summary_rev
```

### Plot

```{r}
plot_districts(districts_rev)
```

