---
title: "Graphs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Graphs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette focuses on all sorts of problems relating to graphs. We will create
a complete graph and an incomplete but connected graph.

```{r setup}
library(lpsugar)
library(ROI)
library(dplyr, warn.conflicts = FALSE)
```

```{r echo=FALSE}
plot_graph <- function(edges) {
    frame()
    par(mai = rep(0.3, 4))
    plot.window(xlim = range(nodes$x), ylim = range(nodes$y), asp = 1)
    
    for (i in 1:nrow(edges)) {
        e <- edges[i, ]
        lines(
            x = c(e$x1, e$x2), 
            y = c(e$y1, e$y2),
            lwd = if (e$active) 4 else 0.5,
            col = if (e$active) "black" else "gray50"
        )
    }
    
    points(nodes$x, nodes$y, pch = 16, cex = 3)
    text(nodes$x, nodes$y, labels = nodes$ind, col = "white", cex = 0.8)
}
```

```{r}
set.seed(125)
n <- 8

nodes <- tibble(
    x = runif(n),
    y = runif(n)
) |> 
    arrange(x) |> 
    mutate(ind = 1:n)

edges_complete <- expand.grid(from = nodes$ind, to = nodes$ind) |> 
    as_tibble() |> 
    filter(from != to) |> 
    mutate(
        x1 = nodes$x[from],
        x2 = nodes$x[to],
        y1 = nodes$y[from],
        y2 = nodes$y[to],
        dx = x1 - x2,
        dy = y1 - y2,
        cost = sqrt(dx^2 + dy^2),
        dx = NULL, dy = NULL,
        active = FALSE,
    )

edges_connected <- edges_complete |> 
    slice_sample(prop = 0.25, weight_by = 1/cost)

plot_graph(edges_complete)
plot_graph(edges_connected)
```


## Minimum Spanning Tree

### Complete graph

We start by creating the cost matrix. It's a symetrical square matrix indicating the
distance from each to each node.

```{r}
cost <- nodes |> 
    select(x, y) |> 
    dist("euclidean") |> 
    as.matrix()
```

We will solve it as a flow problem. We choose an arbitrary root node and make it
send $n$ units of flow, where $n$ is the number of nodes. Each node demands 1 unit.

Our decision variable `flow[i, j]` indicates how many units flow from node `i` to `j`.
Since `flow[i, i]` will be zero, we can fix an upper bound of 0 for the diagonal of `flow`.

```{r}
upper_bound <- matrix(Inf, nrow = n, ncol = n)
diag(upper_bound) <- 0

msp <- lp_problem() |> 
    lp_variable(flow[nodes$ind, nodes$ind], lower = 0, upper = upper_bound)
```

```{r eval=FALSE}
# Alternatively, but slower
msp <- lp_problem() |> 
    lp_variable(flow[nodes$ind, nodes$ind], lower = 0) |> 
    lp_constraint(diag(flow) == 0)
```

The next step is to set our objective. But for this we will need an auxiliary binary variable,
telling us whether an edge has a flow. We call this variable `has_flow` and define it
as binary. Then we connect it to our `flow` variable.

```{r}
msp <- msp |> 
    lp_variable(has_flow[nodes$ind, nodes$ind], binary = TRUE) |> 
    lp_constraint(flow <= n * has_flow)
```

We can now set the objective.

```{r}
msp <- msp |> 
    lp_minimize(sum(cost * has_flow))
```


Now let's restrict the flow of the rest of the nodes. Let's choose the 1st node as our root.
It will have a supply of $n$ units. Each other node will have a supply of $0$ and a demand
of $n$.

```{r}
msp <- msp |> 
    lp_constraint(
        root = sum(flow[1, ]) <= n,
        demand = for (i in nodes$ind) 
            sum(flow[, i]) - sum(flow[i, ]) == 1
    )
```

The full code of the problem is this:

```{r}
upper_bound <- matrix(Inf, nrow = n, ncol = n)
diag(upper_bound) <- 0

msp <- lp_problem() |> 
    lp_variable(
        flow[nodes$ind, nodes$ind], 
        lower = 0, upper = upper_bound
    ) |> 
    lp_variable(
        has_flow[nodes$ind, nodes$ind], 
        binary = TRUE
    ) |> 
    lp_minimize(
        sum(cost * has_flow)
    ) |> 
    lp_constraint(
        aux = flow <= n * has_flow,
        root = sum(flow[1, ]) <= n,
        demand = for (i in nodes$ind[-1]) 
            sum(flow[, i]) - sum(flow[i, ]) >= 1
    )
```

Let's find the solution.

```{r}
msp_solution <- lp_solve(msp)
msp_solution$objective
msp_solution$variables$has_flow
```
```{r echo=FALSE}
hf <- msp_solution$variables$has_flow

ehf <- edges_complete |> 
    rowwise() |> 
    filter(hf[from, to] == 1) |> 
    mutate(active = TRUE) |> 
    select(from, to, active)

edges_complete |> 
    rows_update(ehf, by = c("from", "to")) |> 
    plot_graph()
```


## Pathfinding

## Traveling Salesman Problem
