---
title: "lpsugar"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{lpsugar}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette will teach you the basic elements of a linear problem and
how to define them with `lpsugar`.

```{r setup}
library(lpsugar)
```

## Quick Example: Transportation Problem

The elements of a transportation problem are these:

- Sets
  - $F$ - `factory`
  - $M$ - `market`
  
- Parameters
  - $s$ - `supply` produced by each `factory`.
  - $d$ - `demand` on each `market`.
  - $c$ - `cost` to transport a unit from each `factory` to each `market`.
  
- Variable
  - $t$ - `transport` products from each `factory` to each `market`.
  
- Objective Function
  - $\min \sum_{f \in F} \sum_{m \in M} (c_{fm} t_{fm})$ - `minimize sum(cost * transport)`
  
- Constraints
  - $\sum_{m\in M}({t_{fm}}) \le s_f \quad \forall f\in F$ - 
  `for (f in factory) sum(transport[f, ]) <= supply[f]`
  - $\sum_{f\in F}({t_{fm}}) \ge d_m \quad \forall m\in M$ - 
  `for (m in market)  sum(transport[, m]) >= demand[m]`

Let's define the problem in `lpsugar`.

```{r}
# Sets
factory <- c("A", "B")
market  <- c("Barcelona", "Paris", "Rome")

# Parameters
supply <- c(A = 20, B = 50)
demand <- c(Barcelona = 30, Paris = 15, Rome = 25)
cost <- c(
    # Bar, Par, Rome
       30,  12,  32,
       16,  18,  40
) |> parameter(factory, market)

# Problem
transportation_problem <- lp_problem() |> 
    lp_variable(
        transport[factory, market],
        lower = 0
    ) |> 
    lp_minimize(
        sum(cost * transport)
    ) |> 
    lp_constraint(
        for (f in factory) sum(transport[f, ]) <= supply[f],
        for (m in market)  sum(transport[, m]) >= demand[m]
    )
```

To solve it, you can use any of `ROI`'s applicable solvers. One option is to
load only the desired solver with `library(ROI.plugin.<solver>)` and another
is to load all of `ROI` with `library(ROI)` and let it choose the solver automatically.

```{r}
library(ROI)
solution <- lp_solve(transportation_problem)
solution$objective
solution$variables
```


## Basic Elements

### Sets

Sets are vectors of indices that a variable or parameter will use. For instance,
we could define a set $S=\{a,b,c\}$ to use for a parameter $\{w_i\}_{i \in S}$
and a variable $\{x_i\}_{i \in S}$.

In `lpsugar` we define sets and parameters as base R objects. Variables are defined
inside a problem.

```{r}
S <- c("a", "b", "c")
w <- c(a = 3, b = 2, c = 6)

p <- lp_problem() |>
    lp_variable(x[S])

w
p$variables$x
```

Note: Sets can be numeric, but they must go from `1:n`, where `n` is the length of the set.
This ensures indexing works as expected (`x[1]` is always equal to `x["1"]`).

### Parameters

Parameters are simply fixed values with named dimensions. It's not necessary to
name their dimensions, but it's a good practice, and it pays off in the long run.

For vectors, I recommend using R's name-value syntax, but it's also possible to use `parameter`.
These are all different alternatives for defining $\{w_i\}_{i \in S}$:

```{r}
S <- c("a", "b", "c")
w <- c(a = 3, b = 2, c = 6)
w <- c(3, 2, 6) |> stats::setNames(S)
w <- c(3, 2, 6) |> parameter(S)
```

For matrices, you can use `parameter` to define them comfortable within the code.
Let's define $\{m_{ij}\}_{i\in S,\ j\in R}$.

```{r}
S <- c("a", "b", "c")
R <- c("A", "B")

m <- c(
    1, 2,
    3, 4,
    5, 6
) |> parameter(S, R)

m
```
Notice this works defining matrices by row (as in `matrix(data, byrow = TRUE)`).
If an object is already a matrix, I recommend setting it's names using `rownames` and `colnames`.

```{r}
m <- matrix(1:6, nrow = 3, byrow = TRUE)
rownames(m) <- S
colnames(m) <- R

m
```
For n-dimensional arrays, `parameter` does not work (by design), so the only option is to
use `dimnames`. We will define a parameter $\{a_{ijk}\}_{i\in S,\ j\in R,\ k\in Q}$

```{r}
S <- c("a", "b", "c")
R <- c("A", "B")
Q <- c("Q1", "Q2")

a <- array(1:12, dim = c(3, 2, 2))
dimnames(a) <- list(S=S, R=R, Q=Q)

a
```


### Variables

Variables are numbers that the solver will change in order to find the
optimal solution. In `lpsugar`, variables can be scalars, vectors, or n-dimensional arrays.
Their dimensions are defined using sets. For instance, a variable
$\{x_{ij}\}_{i \in A,\ j \in B}$ can be defined as `lp_variable(x[A, B])`.

By default, their `lower` and `upper` bounds are `-Inf` to `+Inf`.
It's possible to set a variable to `integer` or `binary`. Let us define
a vector variable $\{x_i \in \mathbb{N}\}_{i \in 1..5}$.

```{r}
p <- lp_problem() |> 
    lp_variable(x[1:5], lower = 0, integer = TRUE)
```


### Objective Function

The objective function is a value that the solver will try to minimize or maximize. 
It is set using `lp_minimize()` or `lp_maximize()`.

Let us minimize the following sum:

$$
\sum_{i \in I}\sum_{j \in J}{s_j x_{ij}}
$$

We'll use `sum_over()` to sum over the indices.

```{r}
I <- c(1, 2, 3)
J <- c("a", "b")

s <- c(a = 2, b = 5)

p <- lp_problem() |> 
    lp_variable(x[I, J]) |> 
    lp_minimize(
        sum_over(i=I, j=J, s[j] * x[i, j])
    )

p$objective$coef
```

Sometimes the goal is to find any feasible solution. This can be done by
using `lp_find_feasible()` instead of `lp_solve()`. Alternatively, the objective
coeficients can be set to 0 by using `lp_minimize(0)`. Then, `lp_solve()` also
returns any feasible solution, with the objective value as 0.

```{r}
pf <- p |> lp_minimize(0)
pf$objective
```


### Constraints
