---
title: "lpsugar"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{lpsugar}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette will teach you the basic elements of a linear problem and
how to define them with `lpsugar`.

```{r setup}
library(lpsugar)
```

## Quick Example: Transportation Problem

The elements of a transportation problem are these:

- Sets
  - $F$ - `factory`
  - $M$ - `market`
  
- Parameters
  - $s$ - `supply` produced by each `factory`.
  - $d$ - `demand` on each `market`.
  - $c$ - `cost` to transport a unit from each `factory` to each `market`.
  
- Variable
  - $t$ - `transport` products from each `factory` to each `market`.
  
- Objective Function
  - $\min \sum_{f \in F} \sum_{m \in M} (c_{fm} t_{fm})$ - `minimize sum(cost * transport)`
  
- Constraints
  - $\sum_{m\in M}({t_{fm}}) \le s_f \quad \forall f\in F$ - 
  `for (f in factory) sum(transport[f, ]) <= supply[f]`
  - $\sum_{f\in F}({t_{fm}}) \ge d_m \quad \forall m\in M$ - 
  `for (m in market)  sum(transport[, m]) >= demand[m]`

Let's define the problem in `lpsugar`.

```{r}
# Sets
factory <- c("A", "B")
market  <- c("Barcelona", "Paris", "Rome")

# Parameters
supply <- c(A = 20, B = 50)
demand <- c(Barcelona = 30, Paris = 15, Rome = 25)
cost <- c(
    # Bar, Par, Rome
       30,  12,  32,
       16,  18,  40
) |> parameter(factory, market)

# Problem
transportation_problem <- lp_problem() |> 
    lp_variable(
        transport[factory, market],
        lower = 0
    ) |> 
    lp_minimize(
        sum(cost * transport)
    ) |> 
    lp_constraint(
        for (f in factory) sum(transport[f, ]) <= supply[f],
        for (m in market)  sum(transport[, m]) >= demand[m]
    )
```

To solve it, you can use any of `ROI`'s applicable solvers. One option is to
load only the desired solver with `library(ROI.plugin.<solver>)` and another
is to load all of `ROI` with `library(ROI)` and let it choose the solver automatically.

```{r}
library(ROI)
solution <- lp_solve(transportation_problem)
solution$objective
solution$variables
```


<!-- ## Basic Elements -->

<!-- ### Sets and Parameters -->

<!-- Sets are vectors of indices that a variable or parameter will use. For instance, -->
<!-- we could define a set $S=\{a,b,c\}$ to use for a parameter $\{w_i\}_{i \in S}$  -->
<!-- and a variable $\{x_i\}_{i \in S}$. -->

<!-- In `lpsugar` we define sets and parameters as base R objects. Variables are defined -->
<!-- inside a problem. -->

<!-- ```{r} -->
<!-- S <- c("a", "b", "c") -->
<!-- w <- c(a = 3, b = 2, c = 6) -->

<!-- p <- lp_problem() |>  -->
<!--     lp_variable(x[S]) -->

<!-- w -->
<!-- p$variables$x -->
<!-- ``` -->


<!-- ### Parameters -->

<!-- Parameters are simply fixed values with named dimensions. It's not necessary to -->
<!-- name their dimensions, but it's a good practice, and it pays off in the long run. -->

<!-- For vectors, I recommend using R's name-value syntax, but it's also possible to use `parameter`. -->
<!-- These are all different alternatives for defining $\{w_i\}_{i \in S}$: -->

<!-- ```{r} -->
<!-- S <- c("a", "b", "c") -->
<!-- w <- c(a = 3, b = 2, c = 6) -->
<!-- w <- c(3, 2, 6) |> stats::setNames(S) -->
<!-- w <- c(3, 2, 6) |> parameter(S) -->
<!-- ``` -->

<!-- For matrices, you can use `parameter` to define them comfortable within the code. -->
<!-- Let's define $\{m_{ij}\}_{i\in S,\ j\in R}$. -->

<!-- ```{r} -->
<!-- S <- c("a", "b", "c") -->
<!-- R <- c("A", "B") -->

<!-- m <- c( -->
<!--     1, 2, -->
<!--     3, 4, -->
<!--     5, 6 -->
<!-- ) |> parameter(S, R) -->

<!-- m -->
<!-- ``` -->
<!-- Notice this works defining matrices by row (as in `matrix(data, byrow = TRUE)`). -->
<!-- If an object is already a matrix, I recommend setting it's names using `rownames` and `colnames`. -->

<!-- ```{r} -->
<!-- m <- matrix(1:6, nrow = 3, byrow = TRUE) -->
<!-- rownames(m) <- S -->
<!-- colnames(m) <- R -->

<!-- m -->
<!-- ``` -->
<!-- For n-dimensional arrays, `parameter` does not work (by design), so the only option is to  -->
<!-- use `dimnames`. We will define a parameter $\{a_{ijk}\}_{i\in S,\ j\in R,\ k\in Q}$ -->

<!-- ```{r} -->
<!-- S <- c("a", "b", "c") -->
<!-- R <- c("A", "B") -->
<!-- Q <- c("Q1", "Q2") -->

<!-- a <- array(1:12, dim = c(3, 2, 2)) -->
<!-- dimnames(a) <- list(S=S, R=R, Q=Q) -->

<!-- a -->
<!-- ``` -->


<!-- ### Variables -->

<!-- Variables are numbers that the solver will change in order to find the -->
<!-- optimal solution. In `lpsugar`, variables can be scalars, vectors, or n-dimensional arrays.  -->

<!-- By default, their `lower` and `upper` bounds are `-Inf` to `+Inf`.  -->
<!-- It's possible to set a variable to `integer` or `binary`. -->

<!-- ### Objective Function -->

<!-- ### Constraints -->
