---
title: "Miscellaneous Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Miscellaneous Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(lpsugar)
library(ROI.plugin.highs)
```

## Browser Problem

Imagine we're the developers of a browser called Ours, and we want to prove that it's
better than other any other browser, based on a set of features.

```{r}
browsers <- c("Ours", "Chrome", "Firefox", "Opera", "Safari")
features <- c("Speed", "Ease of use", "Customizability", 
              "Privacy", "Extensions", "Ad-blocker")
```

For each feature $i$, each browser $b$ will have a rating $R_{ib}$. (These are made up)

```{r echo=FALSE}
# Make up ratings from 0 to 10 for each feature
set.seed(7)
ratings <- parameter(0, features, browsers)
ratings[] <- sample(0:10, size = length(ratings), replace = TRUE)
```

```{r}
ratings
```

The question is how important is each feature? Let's give each feature $i$ an optimal weight $w_i$.
Then, we can calculate a $\text{Score}_b(w)$ for each browser $b$.

$$
\text{Score }_b = \sum_{i\in\text{Features}}{w_i R_{ib}} \qquad \forall b\in\text{Browsers}
$$

Let's try giving all the features the same weight. $w_i=1 \quad \forall i \in \text{Features}$.
We get these scores:

```{r}
scores <- colSums(ratings)
sort(scores, decreasing = TRUE)
```

But this is nonsense! This would mean Ours is one of the worst browsers!

Let's fix this. We'll bias the weights $w$ in favor of Ours. We can do this by
maximizing the difference in scores between Ours (which will have the highest score) 
and the second best score. 

$$
\max_w \{\ \text{Score}_\text{ Ours}(w) - \text{Second Best Score }(w) \ \}
$$

We'll make sure each feature *matters* by having $1 \le w_i \le 5$, and make
them integers for simplicity.

```{r}
browser_problem <- lp_problem() |> 
    lp_variable(weight[features], lower = 1, upper = 5, integer = TRUE) |> 
    lp_variable(score[browsers]) |> 
    lp_variable(second_best_score) |>
    lp_maximize(score["Ours"] - second_best_score) |> 
    lp_subject_to(
        for (b in browsers) {
            score[b] == sum_over(i=features, ratings[i, b] * weight[i])
        },
        for (b in browsers) if (b != "Ours") {
            second_best_score >= score[b]
        }
    ) |> 
    lp_solve(solver = "highs")

optimal_weights <- browser_problem$variables$weight
```

Take a look at the new weights. Now speed and extensions are the most important features, and 
customizability is also relevant. 

```{r echo=FALSE}
data.frame(feature = features, weight = optimal_weights, row.names = NULL)
```

With these new weights, we can calculate the new scores for each browser. These prove
that Ours is the best browser!

```{r}
new_scores <- numeric()
for (b in browsers) 
    new_scores[b] <- sum_over(i=features, ratings[i, b] * optimal_weights[i])

sort(new_scores, decreasing = TRUE)
```
