---
title: "Nonlinear"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Nonlinear}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(lpsugar)
library(ROI.plugin.nloptr)
```

## Constrained Maximum Likelihood

Say we have a list $x$ of vectors $x_i = \{x_{i1}, \dots, x_{im}\}$. Each vector
is generated from a $\text{Normal}(\mu_i, \sigma)$. The catch is that each $\mu$ is greater
than the previous:

$$
x_{ij} \sim \text{Normal}(\mu_i, \sigma) \qquad \forall i\in[1,n],\ j\in[1,m_i]
$$

$$
\mu_i \ge \mu_{i-1} \qquad \forall i \in [2,n]
$$

Let's generate the data.

```{r}
set.seed(101)

n <- 5
m <- runif(n, 4, 8) |> round()

true_mu <- 3 + cumsum(runif(n, 0, 1))
true_sigma <- 7

x <- list()

for (i in 1:n) {
    x[[i]] <- rnorm(m[i], mean = true_mu[i], sd = true_sigma) |> round(2)
}

print(x)
```

We will find $\mu$ and $\sigma$ by maximum likelihood. Here's the function that returns
the loglikelihood. It's arguments *must* be the variables of the problem.

```{r}
loglikelihood_x <- function(mu, sigma) {
    LL <- 0
    
    for (i in 1:n) {
        LL_i <- dnorm(x[[i]], mean = mu[i], sd = sigma, log = TRUE)
        LL <- LL + sum(LL_i)
    }
    
    return(LL)
}
```

Now we solve the problem.

```{r}
constrained_mle <- lp_problem() |> 
    lp_var(mu[1:n]) |> 
    lp_var(sigma, lower = 0) |> 
    lp_max_fun(loglikelihood_x) |> 
    lp_con(
        for (i in 2:n) mu[i] >= mu[i-1]
    ) |> 
    lp_solve(
        solver = "nloptr.cobyla",
        start = list(mu = rep(0, n), sigma = 1)
    )

constrained_mle
```

```{r}
unconstrained_mu <- sapply(x, mean)
constrained_mu <- constrained_mle$variables$mu
```

```{r echo=FALSE}
show_mu <- rbind(
    unconstrained_mu = unconstrained_mu,
    constrained_mu = constrained_mu,
    true_mu = true_mu
)

colnames(show_mu) <- paste0("mu[", seq_along(x), "]")
round(show_mu, 2)
```

Constrained optimization gives a better result.

```{r}
cor(unconstrained_mu, true_mu)
cor(constrained_mu, true_mu)
```

We also get an estimator for $\sigma$.

```{r}
constrained_sigma <- constrained_mle$variables$sigma
```


```{r echo=FALSE}
show_sigma <- rbind(
    constrained_sigma = constrained_sigma,
    true_sigma = true_sigma
)

colnames(show_sigma) <- "sigma"
round(show_sigma, 2)
```

