---
title: "Nonlinear"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Nonlinear}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(lpsugar)
library(ROI.plugin.nloptr)
```

## Constrained Maximum Likelihood

Say we have a list $x$ of vectors $x_i = \{x_{i1}, \dots, x_{im}\}$. Each vector
is generated from a $\text{Normal}(\mu_i, \sigma)$. The catch is this: each $mu$ is greater
than the previous:

$$
\mu_i \ge \mu_{i-1} \qquad \forall i \in [2,n]
$$

Let's generate the data.

```{r}
set.seed(101)

n <- 5
m <- runif(n, 4, 8) |> round()

true_mu <- -0.5 + cumsum(runif(n, 0, 1))
true_sigma <- 7

x <- list()

for (i in 1:n) {
    x[[i]] <- rnorm(m[i], mean = true_mu[i], sd = true_sigma) |> round(2)
}

print(x)
```

We will find $\mu$ and $\sigma$ by maximum likelihood. Here's the function that returns
the loglikelihood. It's arguments *must* be the variables of the problem.

```{r}
loglikelihood <- function(mu, sigma) {
    ll <- 0
    
    for (i in 1:n) {
        lli <- dnorm(x[[i]], mean = mu[i], sd = sigma, log = TRUE)
        ll <- ll + sum(lli)
    }
    
    return(ll)
}
```

Now we solve the problem.

```{r}
constrained_mle <- lp_problem() |> 
    lp_var(mu[1:n]) |> 
    lp_var(sigma, lower = 0) |> 
    lp_max_fun(loglikelihood) |> 
    lp_con(for (i in 2:n) mu[i] >= mu[i-1]) |> 
    lp_solve(
        solver = "nloptr.cobyla",
        start = list(mu = rep(0, n), sigma = 1)
    )

constrained_mle
```

```{r echo=FALSE}
unconstrained_means <- sapply(x, mean)
cbind(
    unconstrained_means = unconstrained_means,
    constrained_mle = constrained_mle$variables$mu,
    true_mu = true_mu
) |> round(2)
```

Constrained optimization gives a better result.

```{r}
cor(unconstrained_means, true_mu)
cor(constrained_mle$variables$mu, true_mu)
```

