---
title: "regression"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{regression}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(lpsugar)
library(ROI)
```

This vignette focuses on different regression techniques that can be solved
using linear models.

$$
y_i = \sum_{j=1}^k{\beta_j x_{ij}} + e_i
$$
We start by generating the data we'll use.

```{r}
set.seed(123)
n <- 50
k <- 3

true_beta <- c(2, 3, -5)

x <- matrix(rpois(n*k, 6), nrow = n, ncol = k)
x[, 1] <- 1

e <- rnorm(n)
y <- (x %*% true_beta) + e

head(x)
head(y)
```


## Least Absolute Deviation

The objective of Least Absolute Deviation (LAD) regression is to minimize the absolute
residuals.

$$
\min{\sum_{i=1}^n{|e_i|}}
$$

The absolute value function $|e|$ is not linear, so we have to separate each $e_i$ into:

$$
e_i = e_i^+ - e_i^-
$$
$$
e_i^+ \ge 0,\ e_i^- \ge 0
$$
Then the problem can be written like this:

$$
\begin{array}{rl}
  \min & \sum_{i=1}^n{e_i^+ + e_i^-} \\
  \text{st} & e_i^+ \ge y_i - \hat{y_i} \\
            & e_i^- \ge \hat{y_i} - y_i \\
            & e_i^+ \ge 0 \\
            & e_i^- \ge 0 \\
  \text{where} & \hat{y_i} = \sum_{j=1}^k{\beta_j x_{ij}}
\end{array}
$$

This works. The objective function attempts to push both $e_i^+$ and $e_i^-$ down,
but the constraints ensure that:

- If $e_i > 0\ \Rightarrow\ e_i^+ = e_i,\ e_i^- = 0$.

- If $e_i < 0\ \Rightarrow\ e_i^+ = 0,\ e_i^- = |e_i|$.

Let's write the problem in `lpsugar`.

```{r}
n <- nrow(x)
k <- ncol(x)

lad <- lp_problem() |> 
    lp_var(beta[1:k, 1]) |> 
    lp_var(e_pos[1:n], lower = 0) |> 
    lp_var(e_neg[1:n], lower = 0) |> 
    lp_min(sum(e_pos + e_neg)) |> 
    lp_alias(yhat = x %*% beta) |> 
    lp_con(
        pos = e_pos >= y - yhat,
        neg = e_neg >= yhat - y
    ) |> 
    lp_solve()

cbind(
    true_beta = true_beta,
    lad_beta = lad$variables$beta[, 1] |> round(2)
)
```

