---
title: "Regression and Classification"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Regression and Classification}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(lpsugar)
library(ROI.plugin.highs)
```

This vignette focuses on different regression techniques that can be solved
using linear models.

$$
y_i = \sum_{j=1}^k{\beta_j x_{ij}} + e_i
$$
We start by generating the data we'll use.

```{r}
set.seed(123)
n <- 50
k <- 3

true_beta <- c(2, 3, -5)

x <- matrix(rpois(n*k, 6), nrow = n, ncol = k)
x[, 1] <- 1
head(x)

e <- rnorm(n)
y <- (x %*% true_beta) + e
head(y)
```


## Least Absolute Deviation

The objective of Least Absolute Deviation (LAD) regression is to minimize the absolute
residuals.

$$
\min{\sum_{i=1}^n{|e_i|}}
$$

The absolute value function $|e|$ is not linear, so we have to separate each $e_i$ into:

$$
e_i = e_i^+ - e_i^-
$$
$$
e_i^+ \ge 0,\ e_i^- \ge 0
$$
Then the problem can be written like this:

$$
\begin{array}{rl}
  \min & \sum_{i=1}^n{e_i^+ + e_i^-} \\
  \text{st } & e_i^+ \ge y_i - \hat{y_i} \\
            & e_i^- \ge \hat{y_i} - y_i \\
            & e_i^+ \ge 0 \\
            & e_i^- \ge 0 \\
  \text{where } & \hat{y_i} = \sum_{j=1}^k{\beta_j x_{ij}}
\end{array}
$$

This works. The objective function attempts to push both $e_i^+$ and $e_i^-$ down,
but the constraints ensure that:

- If $e_i > 0\ \Rightarrow\ e_i^+ = e_i,\ e_i^- = 0$.

- If $e_i < 0\ \Rightarrow\ e_i^+ = 0,\ e_i^- = |e_i|$.

Let's write the problem in `lpsugar`.

```{r}
n <- nrow(x)
k <- ncol(x)

lad <- lp_problem() |> 
    lp_var(beta[1:k]) |> 
    lp_var(e_pos[1:n], lower = 0) |> 
    lp_var(e_neg[1:n], lower = 0) |> 
    lp_min(sum(e_pos + e_neg)) |> 
    lp_alias(yhat = x %*% beta) |> 
    lp_con(
        pos = e_pos >= y - yhat,
        neg = e_neg >= yhat - y
    ) |> 
    lp_solve()
```

The estimated $\hat{\beta}$ is quite similar to the `true_beta` $\beta$.

```{r}
cbind(
    true_beta = true_beta,
    lad_beta = lad$variables$beta |> round(2)
)
```

And the absolute deviation is:

```{r}
lad$objective
```


## Support Vector Machines

Support Vector Machines (SVM) are a binary classification method, that uses a hyperplane
to separate the groups.

$$
x_{ij} \in \mathbb{R} \qquad i \in [1,n],\ j\in [1,k]
$$
$$
y_i \in \{\text{TRUE},\ \text{FALSE}\} \qquad i \in [1,n]
$$

Usually, $y$ is represented as $y_i \in \{-1, +1\}$, but I believe it's easier to understand
as a boolean variable.

### Basic Idea

A hard margin SVM finds a vector $\{w_j\}_{j\in[1,k]}$ and scalar $b$ that represents a hyperplane $h$:

$$
h : \sum_{j=1}^k{(h_{j} w_{j})} + b = 0
$$

Classifications works like this. When the hyperplane expression is positive, $y_i$ is 
predicted `TRUE`, when the expression is negative, $y_i$ is predicted false.

$$
\sum_{j=1}^k{(x_{ij} w_{j})} + b > 0 \Rightarrow y_i=\text{TRUE} \qquad \forall i\in[1,n]
$$
$$
\sum_{j=1}^k{(x_{ij} w_{j})} + b < 0 \Rightarrow y_i=\text{FALSE} \qquad \forall i\in[1,n]
$$

Let us generate some random data.

```{r}
set.seed(101)
n <- 20
k <- 2

true_w <- c(3, -4)
true_b <- 2

x <- rnorm(n*k, mean = 5, sd = 3) |> 
    round(2) |> 
    matrix(n, k)

head(x)

y <- x %*% true_w + true_b > 0
head(y)

plot(x, col = ifelse(y, "orangered", "royalblue"), pch = 16)
```

### Hard Margin

The goal of an SVM is to find the greatest possible margin. This is equivalent to 
minimizing the quadratic sum of $w$ coefficients.

$$
\min_w \sum_{j=1}^k{w_j^2}
$$

Subject to the following constraints:

$$
\sum_{j=1}^k{(x_{ij} w_{j})} + b > +1 \qquad \forall i : y_i = \text{TRUE}
$$
$$
\sum_{j=1}^k{(x_{ij} w_{j})} + b < -1 \qquad \forall i : y_i = \text{FALSE}
$$

```{r}
svm_hard <- lp_problem() |> 
    lp_var(w[1:k]) |> 
    lp_var(b) |> 
    lp_min(sum(w^2)) |> 
    lp_constraint(
        hard_margin = for (i in 1:n) {
            predicted <- sum_over(j = 1:k, x[i,j] * w[j]) + b
            if (y[i]) predicted >= +1
            else      predicted <= -1
        }
    ) |> 
    lp_solve()
```

This is the result. The continuous line represent the hyperplane, and the dashed lines
represent the points where the hyperplane formula is $+1$ and $-1$.

```{r echo=FALSE}
plot(x, col = ifelse(y, "orangered", "royalblue"), pch = 16)

with(svm_hard$variables, {
     abline(a = - b / w[2], b = - w[1] / w[2])
     abline(a = (+1 - b) / w[2], b = - w[1] / w[2], lty = 2)
     abline(a = (-1 - b) / w[2], b = - w[1] / w[2], lty = 2)
})
```

### Soft Margin

Explanation coming soon.

```{r}
C <- 0.6 # Regularization hyperparameter

svm_soft <- lp_problem() |> 
    lp_var(w[1:k]) |> 
    lp_var(b) |> 
    lp_var(slack[1:n], lower = 0) |> 
    lp_min(sum(w^2) + C * sum(slack)) |> 
    lp_constraint(
        soft_margin = for (i in 1:n) {
            predicted <- sum_over(j = 1:k, x[i,j] * w[j]) + b
            if (y[i]) predicted >= +1 - slack[i]
            else      predicted <= -1 + slack[i]
        }
    ) |> 
    lp_solve()
```

```{r echo=FALSE}
plot(x, col = ifelse(y, "orangered", "royalblue"), pch = 16)

with(svm_soft$variables, {
     abline(a = - b / w[2], b = - w[1] / w[2])
     abline(a = (+1 - b) / w[2], b = - w[1] / w[2], lty = 2)
     abline(a = (-1 - b) / w[2], b = - w[1] / w[2], lty = 2)
})
```

